Ellipsoids (v1.0): 3D Magnetic modelling of ellipsoidal bodies

### 2017-07-10 manuscript/manuscript.tex
include the reference soliverez2016

A vast literature about the magnetic
modelling of ellipsoidal bodies was developed in which
are to be found the names of many researchers.
Nevertheless, interest in this subject has not yet died
out, as is evidenced by a list of modern papers in this
field.
Besides, the geoscientific community lacks a free easy-to-use
tool to simulate the magnetic field produced by uniformly
magnetized ellipsoids. Software packages that includes the 
ellipsoid model exist, but all very expensive.
Such a tool could prove to be useful either for teaching
and researching geophysics.

@@

By using the magnetization $\mathbf{M}$ defined by Eq. \ref{eq:M-KH},
the total magnetic field $\mathbf{H}(\mathbf{r})$ (Eq. \ref{eq:H})
can be rewritten as follows:
\begin{equation}
\mathbf{H}(\mathbf{r}) = \mathbf{H}_{0}
+ \mathbf{N}(\mathbf{r}) \, \mathbf{K} \, \mathbf{H}^{\dagger} \: ,
\label{eq:H-M-uniform}
\end{equation}
where $\mathbf{N}(\mathbf{r})$ is a symmetrical matrix whose
$ij$-element $n_{ij}(\mathbf{r})$ is given by
\begin{equation}
n_{ij}(\mathbf{r}) =
\frac{1}{4\pi} \frac{\partial^{2} \, f(\mathbf{r})}
{\partial r_{i} \, \partial r_{j}}
\: , \quad i = 1, 2, 3 \: ,
\quad j = 1, 2, 3 \: ,
\label{eq:nij}
\end{equation}
$r_{1} = x$, $r_{2} = y$, $r_{3} = z$ are the elements of
the position vector $\mathbf{r}$ (Eq. \ref{eq:ellipsoid_surface}),
and
\begin{equation}
f(\mathbf{r}) = \iiint_{\vartheta}
\frac{1}{\| \mathbf{r} - \mathbf{r}^{\prime} \|}
\, dx^{\prime}dy^{\prime}dz^{\prime} \: .
\label{eq:f}
\end{equation}
Notice that the scalar function $f(\mathbf{r})$
(Eq. \ref{eq:f}) is proportional
to the gravitational potential that would be produced by the
ellipsoidal body with volume $\vartheta$ if it had a uniform density
equal to the inverse of the gravitational constant.
It can be shown that the elements $n_{ij}(\mathbf{r})$ are
finite whether $\mathbf{r}$ is a point within or without
the volume $\vartheta$ \citep{peirce1902, webster1904}.
The matrix $\mathbf{N}(\mathbf{r})$ (Eq. \ref{eq:H-M-uniform})
is called \textit{depolarization tensor} \citep{soliverez1981, soliverez2008}.

@@

It is convenient to use $\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$
and $\tilde{f}^{\ddagger}(\tilde{\mathbf{r}})$ to define
the function $\tilde{f}(\tilde{\mathbf{r}})$ evaluated,
respectively, at points
$\tilde{\mathbf{r}}$ inside and outside the volume $\vartheta$ of the
ellipsoidal body.
The scalar function $\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$
is given by
\begin{equation}
\tilde{f}^{\dagger}(\tilde{\mathbf{r}}) = \pi \, abc \,
\int_{0}^{\infty} \left( 1
- \frac{\tilde{x}^{2}}{a^{2} + u}
- \frac{\tilde{y}^{2}}{b^{2} + u}
- \frac{\tilde{z}^{2}}{c^{2} + u} \right)
\frac{1}{R(u)} \, du \: , \quad \tilde{\mathbf{r}} \in \vartheta \: ,
\label{eq:fi-tilde}
\end{equation}
where
\begin{equation}
R(u) = \sqrt{\left( a^{2} + u \right)\left( b^{2} + u \right)\left( c^{2} + u \right)} \: .
\label{eq:R}
\end{equation}
This function represents the gravitational potential
that would be produced by the ellipsoidal body at
points located within its volume $\vartheta$ if it
had a uniform density equal to the inverse of the
gravitational constant.
Notice that, in this case, the gravitational potential
is a quadratic function of the spatial coordinates
$\tilde{x}$, $\tilde{y}$, and $\tilde{z}$, which
supported the \citeauthor{maxwell1873}'s (\citeyear{maxwell1873})
postulate about uniformly magnetized ellipsoids.
In a similar way, the function $\tilde{f}^{\ddagger}(\tilde{\mathbf{r}})$
is given by
\begin{equation}
\tilde{f}^{\ddagger}(\tilde{\mathbf{r}}) = \pi \, abc \,
\int_{\lambda}^{\infty} \left( 1
- \frac{\tilde{x}^{2}}{a^{2} + u}
- \frac{\tilde{y}^{2}}{b^{2} + u}
- \frac{\tilde{z}^{2}}{c^{2} + u} \right)
\frac{1}{R(u)} \, du \: , \quad \tilde{\mathbf{r}} \not\in \vartheta \: ,
\label{eq:fe-tilde}
\end{equation}
where $R(u)$ is defined by Eq. \ref{eq:R} and the
parameter $\lambda$ is defined according to the
ellipsoid type as a function of the spatial coordinates
 $\tilde{x}$, $\tilde{y}$, and $\tilde{z}$ (see Appendix B).
For readers interested in additional information about the
parameter $\lambda$, we recommend \citet[p.~234]{webster1904},
\citet[p.~184]{kellogg1929} and \citet{clark1986}.

>>>>

A vast literature about the magnetic
modelling of ellipsoidal bodies was developed in which
are to be found the names of many researchers.
Nevertheless, interest in this subject has not yet died
out, as is evidenced by a list of modern papers in this
field.
Besides, the geoscientific community lacks a free easy-to-use
tool to simulate the magnetic field produced by uniformly
magnetized ellipsoids.
Such a tool could prove to be useful either for teaching
and researching geophysics.

@@

By using the magnetization $\mathbf{M}$ defined by Eq. \ref{eq:M-KH},
the total magnetic field $\mathbf{H}(\mathbf{r})$ (Eq. \ref{eq:H})
can be rewritten as follows:
\begin{equation}
\mathbf{H}(\mathbf{r}) = \mathbf{H}_{0}
+ \mathbf{N}(\mathbf{r}) \, \mathbf{K} \, \mathbf{H}^{\dagger} \: ,
\label{eq:H-M-uniform}
\end{equation}
where $\mathbf{N}(\mathbf{r})$ is a symmetrical matrix whose
$ij$-element $n_{ij}(\mathbf{r})$ is given by
\begin{equation}
n_{ij}(\mathbf{r}) =
\frac{1}{4\pi} \frac{\partial^{2} \, f(\mathbf{r})}
{\partial r_{i} \, \partial r_{j}}
\: , \quad i = 1, 2, 3 \: ,
\quad j = 1, 2, 3 \: ,
\label{eq:nij}
\end{equation}
$r_{1} = x$, $r_{2} = y$, $r_{3} = z$ are the elements of
the position vector $\mathbf{r}$ (Eq. \ref{eq:ellipsoid_surface}),
and
\begin{equation}
f(\mathbf{r}) = \iiint_{\vartheta}
\frac{1}{\| \mathbf{r} - \mathbf{r}^{\prime} \|}
\, dx^{\prime}dy^{\prime}dz^{\prime} \: .
\label{eq:f}
\end{equation}
Notice that the scalar function $f(\mathbf{r})$
(Eq. \ref{eq:f}) is proportional
to the gravitational potential that would be produced by the
ellipsoidal body with volume $\vartheta$ if it had a uniform density
equal to the inverse of the gravitational constant.
It can be shown that the elements $n_{ij}(\mathbf{r})$ are
finite whether $\mathbf{r}$ is a point within or without
the volume $\vartheta$ \citep{peirce1902, webster1904}.
The matrix $\mathbf{N}(\mathbf{r})$ (Eq. \ref{eq:H-M-uniform})
is called \textit{depolarization tensor} \citep{soliverez1981, soliverez2008, soliverez2016}.

@@

It is convenient to use $\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$
and $\tilde{f}^{\ddagger}(\tilde{\mathbf{r}})$ to define
the function $\tilde{f}(\tilde{\mathbf{r}})$ evaluated,
respectively, at points
$\tilde{\mathbf{r}}$ inside and outside the volume $\vartheta$ of the
ellipsoidal body.
The scalar function $\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$
is given by
\begin{equation}
\tilde{f}^{\dagger}(\tilde{\mathbf{r}}) = \pi \, abc \,
\int_{0}^{\infty} \left( 1
- \frac{\tilde{x}^{2}}{a^{2} + u}
- \frac{\tilde{y}^{2}}{b^{2} + u}
- \frac{\tilde{z}^{2}}{c^{2} + u} \right)
\frac{1}{R(u)} \, du \: , \quad \tilde{\mathbf{r}} \in \vartheta \: ,
\label{eq:fi-tilde}
\end{equation}
where
\begin{equation}
R(u) = \sqrt{\left( a^{2} + u \right)\left( b^{2} + u \right)\left( c^{2} + u \right)} \: .
\label{eq:R}
\end{equation}
This function represents the gravitational potential
that would be produced by the ellipsoidal body at
points located within its volume $\vartheta$ if it
had a uniform density equal to the inverse of the
gravitational constant.
Notice that, in this case, the gravitational potential
is a quadratic function of the spatial coordinates
$\tilde{x}$, $\tilde{y}$, and $\tilde{z}$, which
supported \citeauthor{maxwell1873}'s (\citeyear{maxwell1873})
postulate about uniformly magnetized ellipsoids.
In a similar way, the function $\tilde{f}^{\ddagger}(\tilde{\mathbf{r}})$
is given by
\begin{equation}
\tilde{f}^{\ddagger}(\tilde{\mathbf{r}}) = \pi \, abc \,
\int_{\lambda}^{\infty} \left( 1
- \frac{\tilde{x}^{2}}{a^{2} + u}
- \frac{\tilde{y}^{2}}{b^{2} + u}
- \frac{\tilde{z}^{2}}{c^{2} + u} \right)
\frac{1}{R(u)} \, du \: , \quad \tilde{\mathbf{r}} \not\in \vartheta \: ,
\label{eq:fe-tilde}
\end{equation}
where $R(u)$ is defined by Eq. \ref{eq:R} and the
parameter $\lambda$ is defined according to the
ellipsoid type as a function of the spatial coordinates
 $\tilde{x}$, $\tilde{y}$, and $\tilde{z}$ (see Appendix B).
For readers interested in additional information about the
parameter $\lambda$, we recommend \citet[p.~234]{webster1904},
\citet[p.~184]{kellogg1929} and \citet{clark1986}.
